---
title: "hw01_gapminder data exploration"
output: 
md_document:
    variant: markdown_github
---
### Load data
We will use gapminder for this data exploration assginment
```{r}
library(gapminder)
```

### Exploration of data frames
Here are some ways to get familiar with the data frames

```{r}
head(gapminder)
tail(gapminder)
str(gapminder)
class(gapminder)
ncol(gapminder)
nrow(gapminder)
summary(gapminder)
```
### Explore Variables
Here is a demonstration to select and filter data
```{r}
library(dplyr)
Continent=filter(gapminder, continent=="Asia")
Time=Continent$year
pop=Continent$pop
gdp=Continent$gdpPercap

```

### Visualizing data
This is a basic plot relationship between gdp per cap and life expectancy in Asia
```{r}
plot(lifeExp~gdp,Continent, type="l")
```

Now, let's use ggplot to demonstrate more informative graph with all continents
```{r}
library(ggplot2)
ggplot(gapminder, aes(x=gdpPercap, y=lifeExp, color=continent))+geom_point() # show continents in different color 
```

The points are all clustered with all year (1952-2007) presented. Let's explore one year (e.g. 2002) data to see more clear pattern within each continent.

First, "min" and "max" function can be used to explore data range. For example:
```{r}
min(gapminder$year)
max(gapminder$year)
```

Here we plot relationship between gdp per cap and life expectancy
```{r}
dat=subset(gapminder, year==2002)
ggplot(dat, aes(x=gdpPercap, y=lifeExp, color=continent))+geom_point()+
  xlab("GDP per capita") + 
  ylab("Life expectancy (Years)") + ## change to more readable label
  scale_x_log10() # change to a log scale
```

